---
title: "Replication of Study 'Effects of donation collection methods on donation amount: Nudging donation for the cause and overhead' Suk & Mudita (2022, Psychology & Marketing)"
author: "Fan Yang (fay006@ucsd.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc_depth: 3
editor: 
  markdown: 
    wrap: sentence
---

## Introduction

The study *“Effects of donation collection methods on donation amount: Nudging donation for the cause and overhead”* explores how different donation collection methods can nudge individuals to contribute varying amounts to charitable causes and operational overhead.
Using a choice architecture approach, the study examines whether presenting overhead expenses first or last can mitigate “overhead aversion”, a common reluctance among donors to fund operational costs, as opposed to direct causes.
The authors test three methods across several studies: an allocation method, where donors allocate a set amount between cause and overhead; a cause-first addition, where donors contribute to the cause first; and an overhead-first addition, where donors specify overhead contributions before deciding on cause donations.

In Study 2 of the original paper, Suk and Mudita found that presenting the overhead request before the cause request (overhead-first condition) led to higher total donations compared to the cause-first condition.
The goal of this replication project is to test the reliability of this key finding by replicating the original experiment.
In this setup, participants will engage in a charitable donation scenario where they allocate a fixed amount of money between a cause and overhead, depending on whether they encounter the cause-first or overhead-first sequence.
A key challenge in this study is ensuring that the presentation order truly influences donation behavior across various settings and donor profiles.
If confirmed, these findings could help nonprofits design donation requests that encourage higher overall contributions.

Link to the repository: <https://github.com/ucsd-psych201a/suk2023>

Link to the survey: <https://ucsd-psych201a.github.io/suk2023/website/>

Link to the HTML code: <https://github.com/ucsd-psych201a/suk2023/blob/main/website/index.html>

Link to OSF: <https://osf.io/jzx7e/>

Link to pilot data: <https://github.com/ucsd-psych201a/suk2023/blob/main/data/pilotdata.csv>

## Methods

### Power Analysis

In our replication of Study 2, the sample size was initially determined using the effect sizes observed in Study 1 (*d* = 0.41).
To achieve a statistical power of 80%, we estimated that **a sample size of 188 participants** would be necessary.
This calculation was conducted using **G\*Power**, and it ensures that our replication study will have adequate power to detect a similar effect, assuming the effect size is consistent with that observed in the original study.

### Planned Sample

-   **Sample Size**: The study aims to recruit approximately 188 participants, with about 94 participants per condition (cause-first vs. overhead-first).
-   **Sampling Frame**: Participants will be recruited from Prolific, with eligibility limited to adults aged 18 years and older. There will be no other specific controls for demographics like gender and geographic location.
-   **Termination Rule**: Data collection will end once a minimum of 188 participants is reached.

### Materials

-   **Donation Input Form**: Participants will be presented with a donation form (Same with the Appendix C of the original paper) where they could enter amounts to allocate between two categories: (1) "For the cause" (the direct support for the charitable mission) and (2) "For covering charitable organization's operating expenses" (overhead). The order of these categories will depend on the condition to which participants are assigned (cause-first or overhead-first). The total donation amount was automatically computed based on the amounts inputted for each category, ensuring participants did not exceed the allowed limit.
-   **Satisfaction Scale**: After completing the donation decisions, participants will rate their satisfaction with their donation using a 7-point Likert scale (1 = not satisfied at all, 7 = very satisfied).
-   **Unrelated Survey**: We will include an unrelated survey (content not specified in the original study) about daily routine and lifestyle at before the actual experiment to reduce demand characteristics.

### Procedure

**1. Welcome and Consent**: Participants will be asked for their consent to participate in the study.

**2. Unrelated Survey**: To reduce demand characteristics, we included an unrelated survey on daily routines and lifestyle preferences before the donation task.
This survey serves to mask the study’s true focus on donation behavior by making the experiment appear as a general lifestyle study.
By including questions about sleep, hobbies, and daily activities, we aim to decrease participants' awareness of the study's emphasis on charitable donation choices, thereby minimizing potential biases in their responses.

**3. Explanation of Prize**: Participants are informed about a 20% chance to receive an additional prize, which they can choose to keep or donate.

**4. Information on Charitable Donations**: Participants will receive information on how charitable organizations collect and allocate donations, explaining that donations often go toward both (1) supporting the cause and (2) covering operational costs.
They will also be informed that "some organizations receive donations separately for cause and overhead," to clarify the purpose of each type of donation.
At last, participants will be presented with general information of a donation campaign.

**5. Comprehension Checks**: Two comprehension check questions will ensure participants understand the setup, including how charities use donations (supporting both causes and operational expenses) and the nature of their compensation (a guaranteed amount with a 20% chance to earn an additional amount to keep or donate).

**6.Experimental Conditions**: Two conditions are assigned randomly to the participants.

-   Condition 1 (Cause-First): Participants will first specify how much they wish to donate to the cause, followed by how much to the overhead.

-   Condition 2 (Overhead-First): Participants will first specify how much they wish to donate to the overhead, followed by how much to the cause.

**7.Donation Allocation and Total Calculation**: The donation form will automatically calculate and display the total donation amount based on participants’ allocations to the cause and overhead, ensuring they stay within the allowed limit.
Participants will be reminded that they may donate any portion of the allowed amount and will keep whatever they choose not to donate.

**8. Donation Satisfaction and Experience**: After completing the allocation, participants will rate their satisfaction with their donation decision on a 7-point Likert scale ranging from 1 (not satisfied at all) to 7 (very satisfied).
They will also be asked about their previous donation experience.

**9. Demographic information**: Questions about basic demographic information (age, nationality, gender).

**7. Closing and Compensation**: At the end, participants will be thanked for their participation.
All participants will receive a baseline payment for completing the study.
For those selected in the “20% winners” category, additional compensation will be sent, or remaining funds will be donated according to their indicated choices.

### Analysis Plan

**1. Donation Amount Analysis**: To assess the impact of donation collection method (cause-first vs. overhead-first) on donation behavior, we will conduct regression analyses with total donation amount as the dependent variable, and donation method along with demographic variables (e.g., age and gender) as independent variables.
Separate regressions for cause and overhead donations will further analyze how the collection method influences each allocation independently.
Additionally, an independent samples t-test will compare the donation amounts between the two conditions.

**2. Satisfaction Analysis**: To analyze satisfaction, we will start with a simple regression using satisfaction as the dependent variable and condition (cause-first vs. overhead-first) as the independent variable.
We will then expand to a multiple regression model that includes demographic variables (e.g., age, gender) alongside condition to assess if demographics influence satisfaction.
Finally, a comprehensive model will include amounts donated to cause and overhead as additional predictors.

### Differences from Original Study

Our sample size is determined based on a power analysis using the effect size from Study 1, resulting in a target sample size of 188, different from 143 in the original Study 2.
The original study recruited participants exclusively from the UK, while our study collects data online without geographic restrictions.
Additionally, in the original study, the bonus amount was £6, whereas we are using a bonus of \$8 to account for currency differences and ensure consistency in participant incentives.

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample

Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan

Any differences from what was described as the original plan, or “none”.

## Design Overview

-   **How many factors were manipulated?**

    One factor was manipulated: the order of donation requests (cause-first vs. overhead-first).

-   **How many measures were taken?**

    Four measures were taken: total donation amount, amount donated to cause, amount donated to overhead, and satisfaction with the donation decision.

-   **Did it use a within-participants or between-participants design?**

    The study used a between-participants design, with each participant assigned to one condition (cause-first or overhead-first).

-   **Were measures repeated?**

    No, measures were not repeated.
    Each participant made a single donation decision and rated their satisfaction once.

-   **What would have been the consequence of changing one of these decisions? (e.g. moving from within to between participants). Could it have been either way?**

    If a within-participants design had been used, where participants experienced both conditions, order effects could have biased the results, as participants’ decisions in the second condition might be influenced by their experience in the first.
    It could also make the study’s purpose more apparent, increasing demand characteristics.
    Therefore, a between-participants design is more suitable for this experiment.

-   **Were steps taken to reduce demand characteristics?**

    Yes, an important step taken to reduce demand characteristics was the unrelated survey prior to the actual donation experiment in the original study.
    This initial survey served to mask the study’s true intent, reducing the chances that participants would guess that the research was examining the effect of donation order on donation amounts.

-   **Can you think of any potential confounds to the study?**

    -   Prior Donation Experience: Participants with more experience in charitable giving may have stronger preferences regarding donation allocations, which could influence their responses independently of the experimental condition.
    -   Order of Cause and Overhead in Instructions: If one of the conditions (cause or overhead) is introduced first in the instructions, participants might interpret it as more important, leading them to allocate more funds to that category. This could introduce unintended bias based on the presentation order.

-   **Do you see any limitations to generalizability?**

    -   The online sample may not represent the broader population, especially those who donate regularly in real-world contexts. This limits the study’s ability to generalize findings, as real-life donors might have different motivations and behaviors.
    -   In real life, people typically do not receive money or incentives when making donations, as they did in this study. This financial incentive might have influenced participants' donation behaviors, as they were not actually giving from their own funds.
    -   The study set an upper limit on the donation amount (6 pounds), which is unlike real-life situations where donors can choose any amount they wish to give. Participants might subconsciously use the maximum amount as a reference point, which can shape their perception of what an appropriate donation amount is, potentially influencing their decision-making.

## Results

### Data Preparation

Data will be collected through a custom website, using JsPsych packages integrated into our HTML code to guide participants through the experiment.
The collected data is automatically saved in .csv format on our OSF database.
The process will continue until we have 94 valid participants per condition.

#### Load Relevant Libraries and Functions

Necessary packages (e.g., tidyverse) will be loaded to handle data import, analysis, and visualization.
Key packages include those for reading and analyzing .csv data, conducting regression analyses, and performing t-tests.

```{r}
library(jsonlite)
library(tidyverse)
```

The functions used for data preprocessing:

```{r}
# Larger function to concatenate individual processed files
process_csv_files <- function(file_paths, output_file) {
  
  # Nested function to process a single file
  process_file <- function(file_path) {
    
    # Reads file
    testdata <- read_csv(file_path)
    
    # Shortens file to specified columns
    tdcols <- subset(testdata, select = -c(rt, stimulus, trial_type, plugin_version, question_order))
    
    # Shortens file to specified rows
    tdrows <- tdcols[-c(1:13, 16:17), , drop = FALSE]
    
    # Rearranges elements of cleaned data into one cohesive row
    tdarrange <- data.frame(
      check1 = tdrows$response[1],
      check2 = tdrows$response[2],
      cause = tdrows$causeDonation[3],
      overhead = tdrows$expenseDonation[3],
      total = tdrows$totalDonation[3],
      condition = tdrows$questionOrder[3],
      satis_usual = tdrows$response[4],
      age = tdrows$age[5],
      nationality = tdrows$nationality[5],
      gender = tdrows$response[6],
      comment = tdrows$response[7]
    )
    
    # Parses the satisfaction responses, unnests JSON formatting
    tdparsed <- tdarrange %>%
      mutate(satis_usual = lapply(satis_usual, function(x) {
        tryCatch(fromJSON(x), error = function(e) NA)
        }))
    
    # Unnests the satisfaction/frequency column into two separate columns
    tdsep <- tdparsed %>%
      unnest_wider(col = satis_usual, names_sep = "_")
    
    # Renames columns
    tdclean <- tdsep %>% 
      rename(satisfaction = satis_usual_Q0,
             frequency = satis_usual_Q1)
    
    # Returns cleaned data
    return(tdclean)
  }
  
  # Apply nested function to each file
  processed_files <- lapply(file_paths, process_file)
  
  # Combines files, adds identifying number
  all_data <- bind_rows(processed_files, .id = "file_id")
  
  # Saves the combined data
  write_csv(all_data, output_file)
  
  # Returns combined data
  return(all_data)
}
```

#### Import data

We have established a pipeline from the website to an OSF database using DataPipe, which automatically saves each participant’s responses in .csv format as results come in.
We will download the individual .csv files, clean the data, and combine them into a single .csv file.

Codes for importing data:

```{r, eval=FALSE}
file_paths <- c(example files "../szsyve8hnj.csv", "../l1p5o83rfb.csv", "../2q4qg4qfad.csv")
# Defines filepath for .csv file
output_file <- "/Users/Philip/Documents/CSS204_Final Project_Donation/testpilotdata.csv"
```

#### Data exclusion / filtering

Participants who fail both comprehension checks will be excluded from the analysis to ensure data quality and participant understanding.

#### Prepare data for analysis - create columns etc.

We will create specific columns for the following variables: - Participant ID; - Condition (cause-first or overhead-first); - Demographic Variables: age, gender, nationality; - Satisfaction: recorded on a Likert scale; - Frequency: previous donation experience; - Amount Donated: Separate columns for donations to overhead and cause; - Total Amount Donated: sum of donations to cause and overhead; - Bonus Selection: whether participants were selected to receive the X-dollar bonus

Codes for cleaned data:

```{r, eval=FALSE}
# Runs function and produces .csv file
result <- process_csv_files(file_paths, output_file)

## Cleans JSON formatting
# Function removes all of the special charcters from the data
remove_special_characters <- function(entry) {
  gsub("[^a-zA-Z0-9\\s]", "", entry)
}

# Applies function to all relevant columns
cleanresult <- result %>%
  mutate(check1 = sapply(check1, remove_special_characters)) %>% 
  mutate(check2 = sapply(check2, remove_special_characters)) %>% 
  mutate(condition = sapply(condition, remove_special_characters)) %>% 
  mutate(gender = sapply(gender, remove_special_characters)) %>% 
  mutate(comment = sapply(comment, remove_special_characters))

# Edits data for legibility
cleanresult <- cleanresult %>%
  mutate(
    condition = recode(
      condition,
      ForthecauseForcoveringcharitableorganizationsoperatingexpense = 'causefirst',
      ForcoveringcharitableorganizationsoperatingexpenseForthecause = 'overfirst'
    ),
    check1 = gsub("chanceresponse", "", check1),
    check2 = gsub("donationuseresponse", "", check2),
    gender = gsub("gender", "", gender),
    comment = gsub("Q0", "", comment)
  )

# Produces .csv file 
write.csv(cleanresult,"/Users/Philip/Documents/CSS204_Final Project_Donation/pilotdata.csv", row.names = FALSE)
```

### Confirmatory analysis

The confirmatory analysis will focus on testing whether the order of donation requests significantly affects the total amount donated.
We will conduct an independent samples t-test comparing the mean total donation amounts between the two conditions: cause-first and overhead-first.
Additionally, a series of regression analyses will be conducted to further investigate the effect of donation order while controlling for demographic covariates, such as age and gender.
Together, these analyses address the hypothesis that the overhead-first condition will lead to higher total donations.

Using the pilot data of 16 samples we collected, we performed a multiple regression analysis on the total donation amount to establish the analysis structure:

```{r}
cleanresult <- read.csv("/Users/Philip/Documents/CSS204_Final Project_Donation/pilotdata.csv")
# Test regression analysis (total donations ~ other variables)
output <- lm('total ~ condition + satisfaction + frequency + age + nationality + gender', data = cleanresult)
summary(output)
```

```{r}
outputmean <- cleanresult %>% group_by(condition) %>%  summarize(mean = mean(total))
outputsd <- cleanresult %>% summarize(sd = sd(total))
CohensD <- (outputmean$mean[1] - outputmean$mean[2])/outputsd$sd[1]
CohensD
```

### Exploratory analyses

Any follow-up analyses desired (not required).

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.
None of these need to be long.
